### 2014-11-11 bounds-files/conclusion.tex

We have presented a unified framework for bounding the runtimes of dual-tree
algorithms that use cover trees and the standard cover tree pruning dual-tree
traversal (Algorithm \ref{alg:cover-tree-dual}).  Our main result, Theorem
\ref{thm:ct-runtime}, allows plug-and-play runtime bounding of these algorithms.
We have shown that Theorem \ref{thm:ct-runtime} is useful for bounding the
runtime of nearest neighbor search (Theorem \ref{thm:nns}), approximate kernel
density estimation (Theorem \ref{thm:kde-bound}), exact range count, and exact
range search (Theorem \ref{thm:rs}).  With our contribution, bounding a cover
tree dual-tree algorithm is straightforward and only involves bounding the
maximum size of the reference set, $| R^* |$.