### 2014-11-11 bounds-files/sec03-approx-kde.tex

Ram et.~al. present a clever technique for bounding the
running time of approximate kernel density estimation based on the properties of
the kernel, when the kernel is shift-invariant and satisfies a few assumptions
\cite{ram2009}.  We will restate these assumptions and provide an adapted proof
using Theorem \ref{thm:ct-runtime}, which gives a tighter bound.

Approximate kernel density estimation is a common application of dual-tree
algorithms \cite{gray2003nonparametric, nbody}.  Given a query set $S_q$, a
reference set $S_r$ of size $N$, and a kernel function $\mathcal{K}(\cdot,
\cdot)$, the goal is to calculate, for each $p_q \in S_q$, a density estimate
$f(p_q)$ such that $| f(p_q) - f^*(p_q) | < N \epsilon$ for some
parameter $\epsilon$ and

%\vspace*{-0.5em}
\begin{equation}
f^*(p_q) = \sum_{p_r \in S_r} \mathcal{K}(p_q, p_r).
\end{equation}
%\vspace*{-0.5em}

This is related to the well-studied problem of kernel summation, which can
also be solved with dual-tree algorithms \cite{lee2006faster, lee2008fast}.  In
both of those problems, simple geometric observations can be made to accelerate
computation: when $\mathcal{K}(\cdot, \cdot)$ is shift-invariant, faraway points
have very small kernel evaluations.  Thus, trees can be built on $S_q$ and
$S_r$, and node combinations can be pruned when the nodes are far apart while
still obeying the error bound $N \epsilon$.\footnote{Here we consider
only absolute-value approximation.  Our methods can be easily adapted to other
schemes, such as relative-value approximation.  \cite{ram2009} considers
relative-value approximation in far more detail.}

A tree-independent algorithm for solving approximate kernel density estimation
is given as a \texttt{BaseCase()} function in Algorithm \ref{alg:kde_base_case}
and a \texttt{Score()} function in Algorithm \ref{alg:kde_score} (a correctness
proof can be found in \cite{curtin2013tree}).  This function
assumes a shift-invariant kernel $\mathcal{K}(p_i, p_j) = \mathcal{K}(d( p_i,
p_j ))$.  The list $f_p$ holds partial kernel density estimates for
each query point, and the list $f_n$ holds partial kernel density
estimates for each query node.  At the beginning of the dual-tree traversal, the
lists $f_p$ and $f_n$, which are both of size $O(N)$, are each
initialized to 0.

After the dual-tree traversal, the actual kernel density estimates $f$
must be extracted.  This can be done by traversing the query tree and
calculating $f(p_q) = f_p(p_q) + \sum_{\mathscr{N}_i \in T}
f_n(\mathscr{N}_i)$, where $T$ is the set of nodes in $\mathscr{T}_q$ that
have $p_q$ as a descendant.  %It is possible to implement this in a way where
Each query node needs to be visited only once.

Note that this version is far simpler than the algorithm proposed in,
e.g., \cite{gray2003nonparametric}; however, it is sufficient for our runtime
analysis.  Real-world implementations, such as the one found in \textbf{mlpack}
\cite{mlpack2013}, tend to be far more complex.

\begin{algorithm}[tb]
  \begin{algorithmic}[1]
    \STATE {\bfseries Input:} query point $p_q$, reference point $p_r$, list of
  kernel point estimates $\hat{f}_p$
    \STATE {\bfseries Output:} kernel value $\mathcal{K}(p_q, p_r)$

    \medskip
    \STATE $f_p(p_q) \gets f_p(p_q) + \mathcal{K}(p_q, p_r)$
    \RETURN $\mathcal{K}(p_q, p_r)$
  \end{algorithmic}

  \caption{Approximate kernel density estimation \texttt{BaseCase()}}
  \label{alg:kde_base_case}
\end{algorithm}

\begin{algorithm}[tb]
  \begin{algorithmic}[1]
    \STATE {\bfseries Input:} query node $\mathscr{N}_q$, reference node
$\mathscr{N}_r$, list of node kernel estimates $\hat{f}_n$
    \STATE {\bfseries Output:} a score for the node combination $(\mathscr{N}_q,
\mathscr{N}_r)$, or $\infty$ if the combination should be pruned

    \medskip

    \IF{$\mathcal{K}(d_{\min}(\mathscr{N}_q, \mathscr{N}_r)) -
\mathcal{K}(d_{\max}(\mathscr{N}_q, \mathscr{N}_r)) < \epsilon$}
      \STATE $f_n(\mathscr{N}_q) \gets f_n(\mathscr{N}_q) + |
\mathscr{D}^p(\mathscr{N}_r) | \left(\mathcal{K}(d_{\min}(\mathscr{N}_q,
\mathscr{N}_r)) + \mathcal{K}(d_{\max}(\mathscr{N}_q, \mathscr{N}_r))\right) /\;2$
      \RETURN $\infty$
    \ENDIF

    \RETURN $\mathcal{K}(d_{\min}(\mathscr{N}_q, \mathscr{N}_r)) -
\mathcal{K}(d_{\max}(\mathscr{N}_q, \mathscr{N}_r))$
  \end{algorithmic}

  \caption{Approximate kernel density estimation \texttt{Score()}}
  \label{alg:kde_score}
\end{algorithm}

\begin{thm}
Assume that $\mathcal{K}(p_i, p_j) = \mathcal{K}\left(d(p_i, p_j)\right)$ is a shift-invariant kernel with some bandwidth $h > 0$, where $\mathcal{K}(x)$ is smooth, non-negative, monotonically decreasing, and concave for $x \in [0, h]$ and convex for $x \in (h, \infty)$. Then, given a query set $S_q$ and a reference set $S_r$ with
expansion constant $c_r$, an inverse degree of bichromaticity $\nu$ and using the
approximate kernel density estimation \texttt{BaseCase()} and \texttt{Score()}
as given in Algorithms \ref{alg:kde_base_case} and \ref{alg:kde_score},
respectively, with the traversal given in Algorithm \ref{alg:cover-tree-dual},
the running time of approximate kernel density estimation for some $\epsilon$
error parameter is bounded by
%\begin{equation}
$O(c_r^{8 + \log_2 \zeta} \nu N)$
%\end{equation}
where $\zeta = -\mathcal{K}'(h) \mathcal{K}^{-1}(\epsilon) \epsilon^{-1}$.

\label{thm:kde-bound}
\end{thm}

\begin{proof}
It is clear that \texttt{BaseCase()} and \texttt{Score()} both take $O(1)$ time,
so Theorem \ref{thm:ct-runtime} implies the total runtime of the dual-tree
algorithm is bounded by $O(c_r^4 \nu |R^*| N)$.  Thus, we will bound $|R^*|$
using techniques related to those used in \cite{ram2009}.  The bounding of
$|R^*|$ is split into two sections: first, show that when the scale $s_r^{\max}$
is small enough, $R^*$ is empty.  Second, bound $R^*$ when $s_r^{\max}$ is larger.

The \texttt{Score()} function is such that any node in $|R^*|$ for a given query
node $\mathscr{N}_q$ obeys

\vspace*{-0.8em}
\begin{equation}
\mathcal{K}(d_{\min}(\mathscr{N}_q, \mathscr{N}_r)) -
\mathcal{K}(d_{\max}(\mathscr{N}_q, \mathscr{N}_r))
\ge \epsilon.
\end{equation}

Thus, we are interested in the maximum possible value $\mathcal{K}(a) -
\mathcal{K}(b)$ for a fixed value of $b - a > 0$.  Due to the nature of
$\mathcal{K}(\cdot)$, the maximum value of $\mathcal{K}'(\cdot)$ is
$\mathcal{K}'(h)$; therefore, the maximum possible value of $\mathcal{K}(a) -
\mathcal{K}(b)$ is when the interval $[a, b]$ is centered on $h$.  Further,
$\mathcal{K}(a) - \mathcal{K}(b) \le \epsilon$ when $(b - a) \le (-\epsilon /
\mathcal{K}'(h))$.  Note that

\vspace*{-1.5em}
\begin{equation}
d_{\max}(\mathscr{N}_q, \mathscr{N}_r) - d_{\min}(\mathscr{N}_q, \mathscr{N}_r)
= d(p_q, p_r) + 2^{s_r^{\max} + 1} - d(p_q, p_r) + 2^{s_r^{\max} + 1} = 2^{s_r^{\max} + 2}.
\end{equation}
\vspace*{-0.8em}

Therefore, $|R^*| = \emptyset$ when
%\begin{eqnarray}
$2^{s_r^{\max} + 2} \le -\epsilon / \mathcal{K}'(h)$, or when
%\begin{equation}
$s_r^{\max} \le \log_2( -\epsilon / \mathcal{K}'(h) ) - 2$.
%\end{equation}
%
Consider, then, the case when $s_r^{\max} > \log_2( -\epsilon /
\mathcal{K}'(h) ) - 2$.  It is trivial to see that for any $\mathscr{N}_r \in
R^*$, $\mathcal{K}(d_{\max}(\mathscr{N}_q, \mathscr{N}_r)) > \epsilon$;
alternately stated, this means that
%\begin{equation}
$d(p_q, p_r) < \mathcal{K}^{-1}(\epsilon) + 2^{s_r^{\max} + 1}$.
%\end{equation}
Therefore, bounding the number of points in the set
%\begin{equation}
$B_{S_r}(p_q, \mathcal{K}^{-1}(\epsilon) + 2^{s_r^{\max} + 1}) \cap
C_{s_r^{\max}}$
%\end{equation}
is sufficient to bound $|R^*|$.  For notational convenience, define $\omega = \mathcal{K}^{-1}(\epsilon) /
2^{s_r^{\max} + 1} + 1$.  Using Lemma \ref{lem:packing} with $\delta =
2^{s_r^{\max}}$ and $\rho = 2 \omega$ gives $|R^*| = c_r^{3 + \log_2 \omega}$.

%The procedure to bound this will be the same as
%for the nearest neighbor search proof---a packing bound.  See that

%\vspace*{-1.0em}
%\begin{eqnarray}
%B_{S_r}\left(p_q, \mathcal{K}^{-1}(\epsilon) + 2^{s_r^{\max} + 1}\right) &\subseteq&
%B_{S_r}\left(p_r, 2 (\mathcal{K}^{-1}(\epsilon) + 2^{s_r^{\max} + 1})\right) \\
% &=& B_{S_r}\left(p_r, 2^{s_r^{\max} + 2} \left(
%\frac{\mathcal{K}^{-1}(\epsilon)}{2^{s_r^{\max} + 1}} + 1\right)\right).
%\label{eqn:kde_ball}
%\end{eqnarray}
%\vspace*{-0.5em}

%For notational convenience, define $\omega = \mathcal{K}^{-1}(\epsilon) /
%2^{s_r^{\max} + 1} + 1$.  Applying the definition of expansion constant,

%\begin{equation}
%\left| B_{S_r}\left(p_r, 2^{s_r^{\max} + 2} \omega \right) \right|
%\le c_r^{3 + \log_2 \omega} \left|
%B_{S_r}\left(p_r, 2^{s_r^{\max} - 1}\right) \right|.
%\end{equation}

%The total number of nodes in $R^*$ is bounded by the number of disjoint balls of
%size $2^{s_r^{\max} - 1}$ that can be packed into the ball in Equation
%\ref{eqn:kde_ball}.  In the worst case, this packing is perfect, and

%\begin{equation}
%\left| B_{S_r}(p_q, \mathcal{K}^{-1}(\epsilon) + 2^{s_r^{\max} + 1}) \cap
%C_{s_r^{\max}} \right| \le \frac{\left| B_{S_r}\left(p_r, 2^{s_r^{\max} + 2}
%\omega \right) \right|}{\left| B_{S_r}\left(p_r, 2^{s_r^{\max} - 1}\right) \right|} \le c_r^{3
%+ \log_2 \omega}.
%\end{equation}

The value $\omega$ is maximized when $s_r^{\max}$ is minimized.  Using the lower
bound on $s_r^{\max}$, $\omega$ is bounded as
%
%\begin{equation}
%$\omega = \frac{\mathcal{K}^{-1}(\epsilon)}{2^{s_r^{\max} + 1} + 1} \le
%\frac{\mathcal{K}^{-1}(\epsilon)}{2^{s_r^{\max} + 1}} = -2 \mathcal{K}'(h)
%\mathcal{K}^{-1}(\epsilon) \epsilon^{-1}.
%\end{equation}
$\omega = -2 \mathcal{K}'(h) \mathcal{K}^{-1}(\epsilon) \epsilon^{-1}$.
%
Finally, with $\zeta = -\mathcal{K}'(h) \mathcal{K}^{-1}(\epsilon)
\epsilon^{-1}$, we are able to conclude that $|R^*| \le c_r^{3 + \log_2 (2
\zeta)} = c_r^{4 + \log_2 \zeta}$.  Therefore, the entire dual-tree traversal
takes $O(c_r^{8 + \log_2 \zeta} \nu N)$ time.

The postprocessing step to extract the estimates $f(\cdot)$ requires one
traversal of the tree $\mathscr{T}_r$; the tree has $O(N)$ nodes, so this takes
only $O(N)$ time.
%However, there is one more component to consider: the extraction of the
%estimates $f(p_q)$ for each $p_q$.  As discussed earlier, this process can be
%completed in one breadth-first or depth-first traversal of the query tree
%$\mathscr{T}_q$; this takes $O(N)$ time because there are $O(N)$ nodes.%: at
%each node $\mathscr{N}_q$, distribute the partial kernel density estimate to
%each of its children; for each $\mathscr{N}_c \in \mathscr{C}(\mathscr{N}_q)$,
%set
%\begin{equation}
%$f_n(\mathscr{N}_c) \gets f_n(\mathscr{N}_c) +
%f_n(\mathscr{N}_q).$
%\end{equation}
%Continue this process until the leaf nodes are reached; then, for each leaf node
%$\mathscr{N}_l$ which holds point $p_l$, take
%\begin{equation}
%$f(p_l) \gets f_p(p_l) + f_n(\mathscr{N}_l)$
%\end{equation}
%and this will produce the correct estimates $f(\cdot)$.  Because
%there are $O(N)$ nodes and each node has up to $c^4$ children, this process can
%take no longer than $O(c^4 N)$ time.
This is less than the runtime of the
dual-tree traversal, so the runtime of the dual-tree traversal dominates the
algorithm's runtime, and the theorem holds.
\end{proof}

%Define the inverse of the kernel function $\mathcal{K}(\cdot)$ as
%$\mathcal{K}^{-1}(\cdot)$, and define the set $P^*$ as the points held in each
%node in $R^*$. The set $P^*$ for a query node $\mathscr{N}_q$ with
%point $p_q$ can be split into three disjoint sets:

%\begin{eqnarray}
%P^*_l &=& \{ p_r \in P^* : d(p_q, p_r) < h - 2^{s_r^{\max} + 1} \}, \\
%P^*_m &=& \{ p_r \in P^* : h - 2^{s_r^{\max} + 1} \le d(p_q, p_r) \le h +
%2^{s_r^{\max} + 1} \}, \\
%P^*_r &=& \{ p_r \in P^* : d(p_q, p_r) > h + 2^{s_r^{\max} + 1} \}.
%\end{eqnarray}

%First, consider the set $P^*_l$.  For any $p_r \in P^*_l$, we can use the
%concavity of $\mathcal{K}(\cdot)$ in the region $[0, h]$ to see that

%\begin{eqnarray}
%\epsilon &<& \mathcal{K}(\max(0, (d(p_q, p_r) - 2^{s_r^{\max} + 1}))) -
%\mathcal{K}(d(p_q, p_r) + 2^{s_r^{\max} + 1}) \\
% &\le& \mathcal{K}(d(p_q, p_r) + 2^{s_r^{\max} + 1}) - 2^{s_r^{\max} + 2}
%\mathcal{K}'(d(p_q, p_r) + 2^{s_r^{\max} + 1}) - \mathcal{K}(d(p_q, p_r) +
%2^{s_r^{\max} + 1}) \\
% &=& -2^{s_r^{\max} + 2} \mathcal{K}'(d(p_q, p_r) + 2^{s_r^{\max} + 1}).
%\end{eqnarray}

%Denoting the inverse function of $\mathcal{K}'(\cdot)$ as
%$\mathcal{K}'^{-1}_{[a, b]}(\cdot)$ where the output value is restricted to be
%in the range $[a, b]$, we can place a lower bound on $d(p_q, p_r)$:

%\begin{equation}
%d(p_q, p_r) > \mathcal{K}'^{-1}_{[0, h]}\left( \frac{-\epsilon}{2^{s_r^{\max} +
%2}} \right) - 2^{s_r^{\max} + 1}.
%\end{equation}

%Now, consider the set $P^*_u$.  We can make a similar argument using the
%convexity of $\mathcal{K}(\cdot)$ in the region $[h, \infty]$ to see that

%\begin{eqnarray}
%\epsilon &<& \mathcal{K}(d(p_r, p_q) - 2^{s_r^{\max} + 1}) -
%\mathcal{K}(d(p_q, p_r) + 2^{s_r^{\max} + 1}) \\
% &\le& \mathcal{K}(d(p_q, p_r) - 2^{s_r^{\max} + 1}) - \left(\mathcal{K}(d(p_q,
%p_r) - 2^{s_r^{\max} + 1}) + 2^{s_r^{\max} + 2} \mathcal{K}'(d(p_q, p_r) -
%2^{s_r^{\max} + 1}) \\
% &=& -2^{s_r^{\max} + 2} \mathcal{K}'(d(p_q, p_r) - 2^{s_r^{\max} + 1}).
%\end{eqnarray}

%This leads to an upper bound on $d(p_q, p_r)$:

%\begin{equation}
%d(p_q, p_r) < \mathcal{K}'^{-1}_{[h, \infty]}\left(
%\frac{-\epsilon}{2^{s_r^{\max} + 2}}\right) + 2^{s_r^{\max} + 1}.
%\end{equation}

%Lastly, we consider the set $P^*_m$.  Here, we have that

%\begin{eqnarray}
%\epsilon &<& \mathcal{K}(d(p_q, p_r) - 2^{s_r^{\max} + 1}) - \mathcal{K}(d(p_q,
%p_r) + 2^{s_r^{\max} + 1}) \\
% &\le& -2^{s_r^{\max} + 1} \mathcal{K}'(h).
%\end{eqnarray}

%To complete the analysis, we must split into two cases for $s_r^{\max}$.  When
%$s_r^{\max} < \lfloor \log_2 \left

%In the monochromatic case, $\nu = 1$ and the bound is easier to
%understand.
The dependence on $\epsilon$ (through $\zeta$) is expected: as $\epsilon \to 0$
and the search becomes exact, $\zeta$ diverges both because $\epsilon^{-1}$
diverges and also because $\mathcal{K}^{-1}(\epsilon)$ diverges, and the runtime
goes to the worst-case $O(N^2)$; exact kernel density estimation means no nodes
can be pruned at all.

%As we will show, many kernels have a property such that $-\mathcal{K}'(h)
%\mathcal{K}^{-1}(\epsilon)$ has zero dependence on the bandwidth of the kernel.
%Although this result may seem counterintuitive, the only quantities of the
%kernel that affect the runtime are the maximum derivative (which is at the
%inflection point) and the radius of the ball for which $\mathcal{K}(\cdot) \ge
%\epsilon$.

%To see an example of the lack of dependence on kernel bandwidth, consider the
%common Gaussian kernel:

For the Gaussian kernel with bandwidth $\sigma$ defined by $\mathcal{K}_g(d) =
\exp(-d^2 / (2 \sigma^2))$, $\zeta$ does not
depend on the kernel bandwidth; only
the approximation parameter $\epsilon$.  For this kernel, $h = \sigma$ and
therefore $-\mathcal{K}'_g(h) = \sigma^{-1} e^{-1 / 2}$.  Additionally,
$\mathcal{K}_g^{-1}(\epsilon) = \sigma \sqrt{2 \ln (1 / \epsilon)}$.  This means
that for the Gaussian kernel, $\zeta = \sqrt{(-2 \ln \epsilon) / (e
\epsilon^2)}$.  Again, as $\epsilon \to 0$, the
runtime diverges; however, note that there is no dependence on the kernel
bandwidth $\sigma$.  For a reasonably chosen $\epsilon = 0.05$, the runtime is
approximately $O(c_r^{8.89} \nu N)$; for $\epsilon = 0.01$, the runtime is
approximately $O(c_r^{11.52} \nu N)$.  For very small $\epsilon = 0.00001$, the
runtime is approximately $O(c_r^{22.15} \nu N)$.

%Next, consider this sigmoidal kernel: $\mathcal{K}_s(d) = (1 + \exp(\alpha (d -
%\beta)))^{-1}$; note that this is not the more common `sigmoid kernel', which is
%actually a hyperbolic tangent.  Instead, this kernel resembles a step function.
%Some elementary calculus gives an inflection point $h = \beta$, $\mathcal{K}'(h)
%= -\alpha / 4$, and $\mathcal{K}^{-1}(\epsilon) = \ln(\epsilon^{-1} - 1) /
%\alpha + \beta$.  This gives the following expression for $\zeta$:

%\vspace*{-0.8em}
%\begin{equation}
%\zeta = (1/4) ( \epsilon^{-1} \ln(\epsilon^{-1} - 1) +
%(\beta \alpha) / \epsilon ).
%\end{equation}
%\vspace*{-0.8em}
%
%Now, consider the problem of range count, where the number of points in the
%range $[0, u]$ for each query point $p_q \in S_q$ is desired.  If the sigmoidal
%kernel is used with $\beta = u$ for approximate kernel density estimation, then
%as $\alpha \to 0$ and $\epsilon \to 0$, the results will converge to the result
%for exact range count.  Therefore, when $\alpha > 0$ or $\epsilon > 0$, the
%result is approximate range count, and we can use the results of Theorem
%\ref{thm:kde-bound} to show a runtime bound.
%
%Suppose, for the sake of simplicity, that $\alpha = O(\epsilon^{-1})$; then,
%$\zeta = O(\epsilon^{-1} \log \epsilon^{-1} + \epsilon^{-2} u) = O(\epsilon^{-2}
%u)$.  Then, the runtime of approximate range count is bounded as
%
%\vspace*{-1em}
%\begin{eqnarray}
%\begin{equation}
%O(c_r^{8 + \log_2(\epsilon^{-2} u)} \nu N) = O(c_r^{8 + 2 \log_2(\epsilon^{-1}) +
%\log_2(u)} \nu N)
% = O(\epsilon^{-1} u c_r^{8 + 3 \log_2 c_r} \nu N).
%\end{equation}
%\end{eqnarray}
%\vspace*{-1.5em}

%Unfortunately, this bounding technique gives a bad relationship to the
%approximation parameter $\epsilon$ (and rederivations with $\alpha \ne \epsilon$
%suffer from the same issue).  Another drawback of this technique is that it
%cannot account for a range $[l, h]$ with $l > 0$.  Therefore, in the next
%section, we develop an alternate methodology for bounding the running time of
%range search and range count.