### 2014-11-11 bounds-files/sec04-range-search.tex

\begin{algorithm}[tb]
%  \SetAlgoLined
\begin{algorithmic}[1]
    \STATE {\bfseries Input:} query point $p_q$, reference point $p_r$, range
sets $N[p_q]$ and range $[l, u]$
    \STATE {\bfseries Output:} distance $d$ between $p_q$ and $p_r$

%    \medskip
%    \STATE $d \gets \| p_q - p_r \|$
%    \medskip

    \IF{$d(p_q, p_r) \in [r_{\min}, r_{\max}]$ \AND \texttt{BaseCase($p_q$, $p_r$)} not yet called}
    \STATE  $N[p_q] \gets N[p_q] \cup \{ p_r \}$
    \ENDIF

    \RETURN $d$
  \end{algorithmic}
  \caption{Range search \texttt{BaseCase()}}
  \label{alg:rs_bc}
\end{algorithm}

\begin{algorithm}[tb]
  \begin{algorithmic}[1]
    \STATE {\bfseries Input:} query node $\mathscr{N}_q$, reference node
$\mathscr{N}_r$
    \STATE {\bfseries Output:} a score for the node combination $(\mathscr{N}_q,
\mathscr{N}_r)$, or $\infty$ if the combination should be pruned

    \medskip

    \IF{$d_{\min}(\mathscr{N}_q, \mathscr{N}_r) \in [l, u]$ or
$d_{\max}(\mathscr{N}_q, \mathscr{N}_r) \in [l, u]$}
      \RETURN $d_{\min}(\mathscr{N}_q, \mathscr{N}_r)$
    \ENDIF

    \RETURN $\infty$
  \end{algorithmic}
  \caption{Range search \texttt{Score()}}
  \label{alg:rs_sc}
\end{algorithm}

In the range search problem, the task is to find the set of reference points
%\begin{equation}
$N[p_q] = \{ p_r \in S_r : d(p_q, p_r) \in [l, u] \}$
%\end{equation}
for each query point $p_q$, where $[l, u]$ is the given
range.  The range count problem is practically identical, but only the size of
the set, $|N[p_q]|$, is desired.
Our proof works for both of these algorithms
identically, but we will focus on range search.  A \texttt{BaseCase()} and
\texttt{Score()} function are given in Algorithms \ref{alg:rs_bc} and
\ref{alg:rs_sc}, respectively (a correctness proof can be found in
\cite{curtin2013tree}).  The sets $N[p_q]$ (for each $p_q$) are
initialized to $\emptyset$ at the beginning of the traversal.

\begin{thm}
Given a reference set $S_r$ of size $N$ with expansion constant $c_r$, and a
query set $S_q$ of size $O(N)$, a search range of $[l, u]$, and using the range
search \texttt{BaseCase()} and \texttt{Score()} as given in Algorithms
\ref{alg:rs_bc} and \ref{alg:rs_sc}, respectively, with the standard cover tree
pruning dual-tree traversal as given in Algorithm \ref{alg:cover-tree-dual}, and
also assuming that for some $\alpha > 0$,

\vspace*{-1em}
\begin{equation}
\left| \{ p_r: p_r \in S_r, (1 - \alpha) l \le d(p_q, p_r) \le (1 + \alpha) u \}
\right| = O(1),
\end{equation}
\vspace*{-1.5em}

\noindent the running time of range search or range count is bounded by
$O(c_r^{8 + \beta} \nu N)$ with
%\begin{equation}
$\beta = \log_2 (1 + \alpha^{-1})$.
%\end{equation}
\label{thm:rs}
\end{thm}

\vspace*{-2.5em}
\begin{proof}
Both \texttt{BaseCase()} (Algorithm \ref{alg:rs_bc}) and \texttt{Score()}
(Algorithm \ref{alg:rs_sc}) take $O(1)$ time.  Therefore, using Lemma
\ref{thm:ct-runtime} we know that the runtime of the algorithm is bounded by
$O(c_r^4 \nu |R^*| N)$. %  Thus, our only task is to bound the maximum size of the
%reference set, $|R^*|$.
By the pruning rule, for a query node
$\mathscr{N}_q$, the reference set $R^*$
is made up of reference nodes $\mathscr{N}_r$ that are within a margin of
$2^{s_q + 1} + 2^{s_r + 1} \le 2^{s_r^{\max} + 2}$ of the range $[l, u]$.  Given
that $p_r$ is the point in $\mathscr{N}_r$,

\vspace*{-1em}
\begin{equation}
%p_r &\in& \left( B_{S_r}(p_q, u + \lambda_q + \lambda_r) \cap
%C_{s_r^{\max}}\right) \setminus \left( B_{S_r}(p_q, l - \lambda_q - \lambda_r)
%\cap C_{s_r^{\max}} \right) \\
p_r \in \left( B_{S_r}(p_q, u + 2^{s_r^{\max} + 2}) \cap
C_{s_r^{\max}}\right)\setminus \left( B_{S_r}(p_q, l - 2^{s_r^{\max} + 2}) \cap
C_{s_r^{\max}} \right). \label{eqn:rsballs}
\end{equation}
\vspace*{-1em}

A bound on the number of elements in this set is a bound on $|R^*|$.  %First,
%consider the case where $u \le 2^{s_r^{\max} + 1}$.  Then,
%\begin{eqnarray}
%|B_{S_r}(p_q, u + 2^{s_r^{\max} + 1})| - |B_{S_r}(p_q, l - 2^{s_r^{\max} + 1})|
%&\le& |B_{S_r}(p_q, 2^{s_r^{\max} + 2})| \\
%&\le& c_{qr}^3 |B_{S_r}(p_q, 2^{s_r^{\max} - 1})| \\
%&\le& c_{qr}^3
%\end{eqnarray}
%\noindent where the last step follows because $|B_{S_r}(p_q, 2^{s_r^{\max} -
%1})| = 1$ due to the separation invariant.
First, consider the case where $u \le \alpha^{-1} 2^{s_r^{\max} + 2}$.  Ignoring
the smaller ball, take $\delta = 2^{s_r^{\max}}$ and $\rho = 4 (1 +
\alpha^{-1})$ and apply Lemma \ref{lem:packing} to produce the bound

\begin{equation}
|R^*| \le c_r^{4 + \log_2(1 + \alpha^{-1})}.
\end{equation}

% , and ignoring
%the smaller inner ball,
% note that for any $p_r \in P$, $d(p_q, p_r) \le (1 +
%\alpha^{-1}) 2^{s_r^{\max} + 2}$.  Therefore,

%\begin{equation}
%|B_{S_r}(p_q, u + 2^{s_r^{\max} + 2}) \cap C_{s_r^{\max}} | \le |B_{S_r}(p_q,
%(1 + \alpha^{-1}) 2^{s_r^{\max} + 2}) \cap C_{s_r^{\max}} |.
%\end{equation}

%In addition, we know that each point in $C_{s_r^{\max}}$ is separated by a
%distance of at least $2^{s_r^{\max}}$.  We use a packing argument similar to the
%one used in Theorem \ref{thm:nns}.

%Given that $P$ is the set of points held in the nodes $R$, it is true that for
%any $p_r \in P$, $d(p_q, p_r) \le (1 + \alpha^{-1}) 2^{s_r^{\max} + 1}$.
%Therefore,

%\vspace*{-1em}
%\begin{eqnarray}
%|B_{S_r}(p_q, (1 + \alpha^{-1}) 2^{s_r^{\max} + 2})| &\le& |B_{S_r}(p_r, (1 +
%\alpha^{-1}) 2^{s_r^{\max} + 3})| \\
%|B_{S_r}(p_q, u + 2^{s_r^{\max} + 1})| - |B_{S_r}(p_q, l - 2^{s_r^{\max} + 1})|
%&\le& |B_{S_r}(p_q, (1 + \alpha^{-1}) 2^{s_r^{\max} + 1})| \nonumber \\
%&\le& |B_{S_r}(p_r, (1 + \alpha^{-1}) 2^{s_r^{\max} + 2})| \nonumber \\
%&\le& |B_{S_r \cup \{ p_q \}}(p_q, (1 + \alpha^{-1}) 2^{s_r^{\max} + 1})
%\nonumber \\
%&\le& c_{qr}^{(2 + \log_2(1 + \alpha^{-1}))} |B_{S_r \cup \{ p_q \}}(p_q, 2^{s_r^{\max} -
%1})| \nonumber \\
%&\le& c_r^{(4 + \log_2(1 + \alpha^{-1}))} | B_{S_r}(p_r, 2^{s_r^{\max} - 1}) |.
%&\le& (1 + \alpha^{-1}) c_r^{3 + \log_2 c_r} | B_{S_r}(p_r, 2^{s_r^{\max} - 1})
%|.
%&\le& (1 + \alpha^{-1}) c_{qr}^{2 + \log_2 c_{qr}}
%\end{eqnarray}
%\vspace*{-1em}

%Because each point in $P \subseteq C_{s_r^{\max}}$ is separated by at least
%$2^{s_r^{\max}}$, $| B_{S_r}(p_r, 2^{s_r^{\max} - 1}) | = 1$ for any $p_r \in
%P$.  Thus, the total number of nodes in $R$ is the number of balls of radius
%$2^{s_r^{\max} - 1}$ that can be packed into a ball of radius $(1 + \alpha^{-1})
%2^{s_r^{\max} + 2}$.  In the worst case, this packing is perfect, giving

%\vspace*{-1em}
%\begin{equation}
%\left| B_{S_r}\left(p_q, (1 + \alpha^{-1}) 2^{s_r^{\max} + 2}\right) \cap C_{s_r^{\max}}
%\right| \le \frac{\left| B_{S_r}\left( p_r, (1 + \alpha^{-1}) 2^{s_r^{\max} +
%3}\right) \right|}{\left| B_{S_r}\left( p_r, (1 + \alpha^{-1}) 2^{s_r^{\max} -
%1}\right) \right|} \le c_r^{4 + \log_2 (1 + \alpha^{-1})}.
%\end{equation}
%\vspace*{-1em}

%Therefore, in this case, $|R^*| \le c_r^{4 + \log_2 (1 + \alpha^{-1})}$.
Now, consider the other case: $u > \alpha^{-1} 2^{s_r^{\max} + 1}$.
%Then,
%\begin{eqnarray}
%B_{S_r}(p_q, u + 2^{s_r^{\max} + 1}) &\subseteq& B_{S_r}(p_q, (1 + \alpha) u), \\
%B_{S_r}(p_q, l - 2^{s_r^{\max} + 1}) &\supseteq& B_{S_r}(p_q, (1 - \alpha) l).
%\end{eqnarray}
This means

\vspace*{-1em}
\begin{equation}
B_{S_r}(p_q, u + 2^{s_r^{\max} + 1}) \setminus B_{S_r}(p_q, l - 2^{s_r^{\max} +
1}) \subseteq B_{S_r}(p_q, (1 + \alpha) u) \setminus B_{S_r}(p_q, (1 - \alpha)
l)
\end{equation}

\noindent and, due to the assumptions, $| B_{S_r}(p_q, (1 + \alpha) u) | - |
B_{S_r}(p_q, (1 - \alpha) l) | = O(1)$, $|R^*| = O(1)$.  Thus, the theorem
holds.
\end{proof}
\vspace*{-0.4em}

%\begin{figure}
%\begin{centering}
%\input{rs-bounds}
%\caption{The shape of geometry to come.}
%\label{fig:rs_shapes}
%\end{centering}
%\end{figure}

%Figure \ref{fig:rs_shapes} details the possibilities when a particular
%combination is visited.

%\begin{enumerate}
%\item If the reference node is entirely contained in the range $[l +
%\lambda_q, u - \lambda_q]$ (as in $\mathscr{N}_{r1}$), then every
%descendant point of the reference node will be in the results for every
%descendant point of the query node; these can be added and then the combination
%can be pruned.\footnote{Because we have assumed that each query point will have
%$O(1)$ results, the total runtime of this step for each query node is $O(1)$, or
%$O(N)$ during the whole algorithm.  We will not consider the runtime
%contribution of this situation further, because other factors will outweigh
%this.}

%\item If the reference node is not contained at all in the range $[l -
%\lambda_q, u + \lambda_q]$ (as in $\mathscr{N}_{r2}$ and
%$\mathscr{N}_{r3}$), then no descendant points of the reference node will be in
%the results for any descendant point of the query node; thus, the combination
%can simply be pruned.

%\item If the reference node overlaps with the ranges $[l - \lambda_q,
%l + \lambda_q]$ or $[u - \lambda_q, u + \lambda_q]$ (as in
%$\mathscr{N}_{r4}$, $\mathscr{N}_{r5}$, and $\mathscr{N}_{r6}$), the node
%combination cannot be pruned by \texttt{Score()}, and must be descended.  This
%is the interesting case that we will focus on hereafter.  Only combinations of
%this type will make up the reference nodes in $|R_r^*|$.
%\end{enumerate}

%In the last situation, a reference node $\mathscr{N}_r$ with center $p_r$ and
%furthest descendant distance $\lambda_r$ satisfies at least one of the following
%two conditions:

%\begin{eqnarray}
%p_r &\in& B_{S_r}(p_q, u + \lambda_q + \lambda_r) \setminus B_{S_r}(p_q, u -
%\lambda_q - \lambda_r), \\
%p_r &\in& B_{S_r}(p_q, l + \lambda_q + \lambda_r) \setminus B_{S_r}(p_q, l -
%\lambda_q - \lambda_r).
%\end{eqnarray}

%Assuming first that $l - \lambda_q - \lambda_r > 0$, the reference node lies in
%the outer shell of a ball.  This is true of any $\mathscr{N}_r \in R_r$ for any
%cover set $R_r$ during the traversal.  If we can bound the size of the set

%\begin{equation}
%\left(B_{S_r}(p_q, u + \lambda_q + \lambda_r) \setminus B_{S_r}(p_q, u -
%\lambda_q - \lambda_r) \right) \cup \left(B_{S_r}(p_q, l + \lambda_q +
%\lambda_r) \setminus B_{S_r}(p_r, l - \lambda_q - \lambda_r) \right)
%\end{equation}

%\noindent then we will have bounded $| R_r^* |$.  First, observe that $\lambda_q
%+ \lambda_r \le 2^{s_r^{\max} + 1}$; then, the following algebraic
%manipulations lead us to our result.

%\begin{cor}
%In the monochromatic case ($S_q = S_r$, and $c_r = c_{qr} = c$), the
%running time of range search or range count is bounded by $O(c^{8 + \beta} N)$.
%\end{cor}

To understand the relation of the running time to the bounding quantity
$\alpha$, consider (for the monochromatic case with $\nu \sim O(1)$) $\alpha =
(1 / 3)$; this yields a running time of $O(c^8 N)$.  $\alpha = (1 / 7)$ yields
$O(c^9 N)$, $\alpha = (1 / 15)$ yields $O(c^{10} N)$, and so forth.  As $\alpha$
gets smaller, the exponent on $c$ gets larger, and diverges as $\alpha \to 0$.

The assumption that on number of points in the range of $[(1 - \alpha)l, (1 +
\alpha)u]$ for a single query is not too restrictive, but must be satisfied for
the result to be meaningful.  After all, if that assumption is not satisfied,
then it is not possible to perform range search in $O(1)$ time for each query
point, making an $O(N)$ bound for the whole query set impossible.

To show where the assumption is valid, consider that when presented with a very
large dataset, a practitioner may perform monochromatic range search on a range
$[l, u]$ for a geometric subset of their data; for instance, a hypersquare
with side length $2u$.  This might be done in order to find a range $[l, u]$
that produces a manageable number of results (if the range is too large, the
number of results returned for each query point may be enormous).  Then, that
practitioner may scale to a larger dataset by simply increasing the size of
their hypersquare.  In this setting, the number of points in the range of
$[(1 - \alpha)l, (1 + \alpha)u]$ for a single query remains $O(1)$ with respect
to $N$.  Alternately, it should be possible to tighten the range $[l, u]$ as $N$
increases in such a way that the number of points in the range remains constant.

%In cases where this assumption is not thought to be valid, it is straightforward
%to prove a bound on the running time of range search or range count by using the
%aspect ratio of the data, as in \cite{curtin2014kernel}.  However, the aspect
%ratio is particularly sensitive to datasets that have arbitrarily close points,
%and it is difficult to make assumptions about the aspect ratio.

Alternately, if the assumption is not valid but it can be shown that the number
of points in the range $[(1 - \alpha)l, (1 + \alpha)u]$ is $O(k)$ where $k$ has
some dependence on $N$, then the running time can be easily bounded as $O(c^{4 +
\beta} kN)$, in the monochromatic case with $\nu \sim O(1)$.

%Also, note that the proof considers points with distance from $p_q$ in
%the range $[l + \lambda_q + \lambda_r, u - \lambda_q - \lambda_r]$; however, the
%pruning rule could also prune points that lie entirely in the range $[l, u]$ for
%additional speedup (Algorithm \ref{alg:rs_sc} does not do this because there is
%no theoretical benefit in our proof).  Unfortunately, the expansion constant
%does not make working with slices of balls easy, and as a result the runtime
%bound as given is likely to be loose when this additional prune is made,
%especially when $u - l$ is large.