### 2014-11-11 bounds.tex

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\hypersetup{colorlinks,
            urlcolor=blue,
            citecolor=green,
            linkcolor=red}
\usepackage{url}
%\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09
\usepackage{mathrsfs}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{color}

\usepackage{pstricks}
\usepackage{tikz}
%\usepackage{subcaption}

\title{Plug-and-play dual-tree algorithm runtime analysis}

%\author{
%David S.~Hippocampus\thanks{ Use footnote for providing further information
%about author (webpage, alternative address)---\emph{not} for acknowledging
%funding agencies.} \\
%Department of Computer Science\\
%Cranberry-Lemon University\\
%Pittsburgh, PA 15213 \\
%\texttt{hippo@cs.cranberry-lemon.edu} \\
%\And
%Coauthor \\
%Affiliation \\
%Address \\
%\texttt{email} \\
%\AND
%Coauthor \\
%Affiliation \\
%Address \\
%\texttt{email} \\
%\And
%Coauthor \\
%Affiliation \\
%Address \\
%\texttt{email} \\
%\And
%Coauthor \\
%Affiliation \\
%Address \\
%\texttt{email} \\
%(if needed)\\
%}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

% Misc. math commands.  I couldn't figure out how to make the Unicode cat
% character, sadly.
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\var}{\operatorname{var}}

\newcommand{\K}{\mathcal{K}}
\newcommand{\hilbert}{\mathcal{H}}
\newcommand{\Real}{\mathcal{R}}

\renewcommand{\algorithmicforall}{\textbf{for each}}
\renewcommand{\algorithmicrepeat}{\textbf{do}}
\renewcommand{\algorithmicuntil}{\textbf{while}}

\newtheorem{defn}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{cor}{Corollary}

%\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle

\vspace*{-0.8em}
\begin{abstract}
%A large collection of machine learning algorithms contain subproblems that can
%be efficiently solved with trees.  In the situation where a dual-tree algorithm
%with cover trees is used to solve a problem, rigorous worst-case runtime bounds
%can often be shown.  A worst-case runtime bound is proven for all cover tree
%dual-tree algorithms; this bound can be employed to give useful runtime bounds
%for individual problems.  We show a new, tighter bound of $O(c^9 \nu N)$ for
%all-nearest-neighbor search and also provide linear runtime bounds for
%approximate kernel density estimation and exact range count/search under some
%mild assumptions.  These are the first linear runtime bounds we are aware of for
%range count/search.
Numerous machine learning algorithms contain pairwise statistical
problems in their core. Often, tree structures are used
to solve these problems efficiently. Dual-tree
algorithms using cover trees have
rigorous worst-case runtime guarantees for some of these problems. In this paper, we present a {\em
problem-independent} runtime guarantee for {\em any} dual-tree
algorithm using the cover tree, separating out the problem-dependent and
the problem-independent elements. This allows us to just plug in
bounds for the problem-dependent elements to get runtime guarantees
for dual-tree algorithms for any pairwise statistical problem. We
demonstrate this plug-and-play procedure for nearest-neighbor search
and approximate kernel density estimation to get improved runtime
guarantees.  Under some assumptions, we also present the first linear runtime
guarantee for dual-tree based range search.
\end{abstract}

\vspace*{-1em}

%\section{Dual-tree algorithms for pairwise statistical problems (Introduction)}
\section{Dual-tree algorithms}
\vspace*{-0.5em}

%\begin{itemize}
%\item What are pairwise statistical problems? Examples?
%
%A surprising number of machine learning algorithms have bottlenecks which, if
%implemented naively, depend on the calculation of potentially large numbers of
%distances.  A common example is nearest neighbor search: if search
%for the nearest neighbor of a single query point in a reference set of size $N$,
%simple linear scan costs $O(N)$ time.  For many decades, spatial data structures
%have been employed to accelerate this type of computation, beginning with the
%quad tree \cite{finkel1974quad}; building trees on the reference set often
%results in expected runtimes of $O(\log N)$ per query point.
%
A surprising number of machine learning algorithms have computational
bottlenecks that can be expressed as pairwise statistical problems.
All-nearest-neighbor search is one such problem, naively requiring $O(N)$ time
to answer in single query in a reference set of size $N$; answering $O(N)$
queries subsequently requiring the prohibitive $O(N^2)$ time. Kernel density
estimation is also a pairwise statistical problem, naively requiring $O(N^2)$
time to answer $O(N)$ queries. The reference set is typically indexed with
spatial data structures to accelerate this type of computation
\cite{finkel1974quad, langford2006}; these result in $O(\log N)$ runtime per
query under favorable conditions.

%\item What are dual-tree algorithms? Brief explanation.
%
%More recently, Gray and Moore \cite{nbody} observed that problems often have
%large query {\em sets}, not just a few query points, and introduced dual-tree
%algorithms.  One example is all-nearest-neighbor search: find the nearest
%neighbor of every point in the reference set.  Simple implementations scale
%poorly, but dual-tree algorithms can obtain orders-of-magnitude speedup for a
%multitude of problems.
%
%Following this, numerous dual-tree algorithms have been developed to solve a
%wide variety of problems; these include kernel density estimation
%\cite{gray2003nonparametric}, fast max-kernel search \cite{curtin2013fast},
%approximate nearest neighbor search \cite{ram2009rank}, mean shift
%\cite{wang2007fast}, minimum spanning tree calculation
%\cite{march2010euclidean}, $n$-point correlation function estimation
%\cite{march2012fast}, kernel summation \cite{lee2006faster}, particle
%smoothing \cite{klaas2006fast}, variational inference \cite{amizadeh2012variational} and range search \cite{nbody}.
%
Gray and Moore \cite{nbody} generalized the fast-multipole-method from
computational physics to obtain {\em dual-tree algorithms}. These are extremely
useful when there are large query {\em sets}, not just a few query points.
%\item What (pairwise statistical) problems have they been applied to?
%
Following this, numerous dual-tree algorithms have been developed to solve a
wide variety of problems; these include kernel density estimation
\cite{gray2003nonparametric}, fast max-kernel search \cite{curtin2013fast},
approximate nearest neighbor search \cite{ram2009rank}, mean shift
\cite{wang2007fast}, minimum spanning tree calculation
\cite{march2010euclidean}, $n$-point correlation function estimation
\cite{march2012fast}, kernel summation \cite{lee2006faster}, particle
smoothing \cite{klaas2006fast}, variational inference \cite{amizadeh2012variational} and range search \cite{nbody}.
%
%\item Existing analyses of dual-tree algorithms
Some of these algorithms are derived using the cover tree
\cite{langford2006}, a data structure with beneficial theoretical qualities.
Dual-tree all nearest-neighbor search and approximate kernel density estimation with cover trees have $O(N)$ runtime guarantees for $O(N)$ queries \cite{ram2009}; minimum spanning tree calculation scales as $O(N \log N)$ \cite{march2010euclidean}. Other problems have similar worst-case guarantees \cite{curtin2014dual, march2013multi}.
% Is there a better reference for the npcf runtime bound?

%\item General tree/problem independent dual-tree algorithms (maybe)
%\item What is lacking -- general dual-tree analysis (since the algorithms are so similar in structure)
%A recent framework for understanding and unifying dual-tree algorithms
%observes that all dual-tree algorithms are particularly similar, and generalizes
%them \cite{curtin2013tree}.  A
%nice benefit of this viewpoint is that it allows quick and easy development of
%new dual-tree algorithms.  In this paper, we make the same observation about
%runtime proofs for cover tree dual-tree algorithms, and show that
%runtime analysis of generalized cover tree dual-tree algorithms is
%possible.  Thus, the contributions of this paper are:
%
%\vspace*{-0.5em}
%\begin{itemize}
%  \item A general runtime bound for dual-tree algorithms that use the cover
%tree.
%
%  \item \mbox{A nearest neighbor search runtime bound,
%slightly improving on existing linear-time results.}
%
%  \item An approximate kernel density estimation runtime bound,
%improving on existing linear-time results.
%
%  \item Novel runtime analysis of dual-tree range count and range
%search.
%\end{itemize}
Curtin et.~al. \cite{curtin2013tree} observe that all dual-tree
algorithms are similar and propose a unified framework which allows the
development of dual-tree algorithms in a problem independent (and tree
independent) manner. In this paper, we make the same observation about runtime
proofs for cover tree dual-tree algorithms, and show that runtime analysis
of a generalized dual-tree algorithm is possible. The specific contributions of
this paper are:
\vspace*{-0.3em}
\begin{itemize}
  \item A general runtime bound for dual-tree algorithms that use the cover
tree.
  \item An improved linear runtime guarantee for all nearest-neighbor search.
  \item An improved linear runtime guarantee for approximate kernel density
estimation.
  \item The first linear runtime guarantee for dual-tree range count and range
search.
\end{itemize}

%The following section will detail theory prerequisites, and then introduce the
%general runtime bound.

%\item This paper's contribution:
%  \begin{itemize}
%  \item Problem independent dual-tree algorithm analysis
%  \item Examples of final problem dependent runtime bounds for NNS and KDE which match existing results (from LTAPS) -- implying nothing is lost yet for doing problem independent analysis
%  \item Novel runtime analyses of dual-tree range count (or range search).
%  \end{itemize}
%\item This paper's structure (to be written last)
%\end{itemize}

\vspace*{-1em}
\section{Preliminaries}
\vspace*{-0.6em}

\input{bounds-files/sec01-supporting-lemmas.tex}

\vspace*{-0.6em}
\section{General runtime bound}
\vspace*{-0.4em}

\input{bounds-files/general-runtime-bound.tex}

\vspace*{-0.8em}
\section{Nearest neighbor search}
\vspace*{-0.4em}

\input{bounds-files/sec02-nearest-neighbor-search.tex}

\vspace*{-0.6em}
\section{Approximate kernel density estimation}
\vspace*{-0.6em}

\input{bounds-files/sec03-approx-kde.tex}

\vspace*{-0.8em}
\section{Range Search and Range Count}
\vspace*{-0.4em}

\input{bounds-files/sec04-range-search.tex}

%\section{Range Search with Aspect Ratio}

%\input{bounds-files/sec05-range-search-aspect-ratio.tex}

\section{Conclusion}

\input{bounds-files/conclusion.tex}

\bibliographystyle{plain}
\bibliography{bounds}

% Appendix doesn't have anything useful or non-straightforward.
%\appendix
%\input{appendix-ec}

\end{document}