### 2014-11-13 figures/balanced-cover-tree.tex
Four figures I made.  I'm not sure if I like graphviz or not.

\begin{tikzpicture}[>=latex,line join=bevel,scale=0.47]
  \pgfsetlinewidth{1bp}
%%
\begin{scope}
  \pgfsetstrokecolor{black}
  \definecolor{strokecol}{rgb}{1.0,1.0,1.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{1.0,1.0,1.0};
  \pgfsetfillcolor{fillcol}
  \filldraw (0bp,0bp) -- (0bp,180bp) -- (486bp,180bp) -- (486bp,0bp) -- cycle;
\end{scope}
\begin{scope}
  \pgfsetstrokecolor{black}
  \definecolor{strokecol}{rgb}{1.0,1.0,1.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{1.0,1.0,1.0};
  \pgfsetfillcolor{fillcol}
  \filldraw (0bp,0bp) -- (0bp,180bp) -- (486bp,180bp) -- (486bp,0bp) -- cycle;
\end{scope}
\begin{scope}
  \pgfsetstrokecolor{black}
  \definecolor{strokecol}{rgb}{1.0,1.0,1.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{1.0,1.0,1.0};
  \pgfsetfillcolor{fillcol}
  \filldraw (0bp,0bp) -- (0bp,180bp) -- (486bp,180bp) -- (486bp,0bp) -- cycle;
\end{scope}
  \pgfsetcolor{black}
  % Edge: c2 -> c6
  \draw [-] (228.43bp,74.834bp) .. controls (218.25bp,64.938bp) and (204.48bp,51.546bp)  .. (185.8bp,33.385bp);
  % Edge: c3 -> c9
  \draw [-] (387bp,71.697bp) .. controls (387bp,63.983bp) and (387bp,54.712bp)  .. (387bp,36.104bp);
  % Edge: c3 -> c10
  \draw [-] (401.57bp,74.834bp) .. controls (411.75bp,64.938bp) and (425.52bp,51.546bp)  .. (444.2bp,33.385bp);
  % Edge: c2 -> c8
  \draw [-] (257.57bp,74.834bp) .. controls (267.75bp,64.938bp) and (281.52bp,51.546bp)  .. (300.2bp,33.385bp);
  % Edge: c1 -> c4
  \draw [-] (84.43bp,74.834bp) .. controls (74.25bp,64.938bp) and (60.476bp,51.546bp)  .. (41.796bp,33.385bp);
  % Edge: c1 -> c5
  \draw [-] (99bp,71.697bp) .. controls (99bp,63.983bp) and (99bp,54.712bp)  .. (99bp,36.104bp);
  % Edge: root -> c1
  \draw [-] (221.75bp,150.67bp) .. controls (197.4bp,138.83bp) and (157.28bp,119.33bp)  .. (120.33bp,101.37bp);
  % Edge: root -> c2
  \draw [-] (243bp,143.7bp) .. controls (243bp,135.98bp) and (243bp,126.71bp)  .. (243bp,108.1bp);
  % Edge: c2 -> c7
  \draw [-] (243bp,71.697bp) .. controls (243bp,63.983bp) and (243bp,54.712bp)  .. (243bp,36.104bp);
  % Edge: root -> c3
  \draw [-] (264.25bp,150.67bp) .. controls (288.6bp,138.83bp) and (328.72bp,119.33bp)  .. (365.67bp,101.37bp);
  % Node: c7
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (243bp,18bp) ellipse (27bp and 18bp);
  \draw (243bp,18bp) node {$\mathscr{N}_h$};
\end{scope}
  % Node: c9
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (387bp,18bp) ellipse (27bp and 18bp);
  \draw (387bp,18bp) node {$\mathscr{N}_j$};
\end{scope}
  % Node: c8
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (315bp,18bp) ellipse (27bp and 18bp);
  \draw (315bp,18bp) node {$\mathscr{N}_i$};
\end{scope}
  % Node: c3
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (387bp,90bp) ellipse (27bp and 18bp);
  \draw (387bp,90bp) node {$\mathscr{N}_d$};
\end{scope}
  % Node: c2
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (243bp,90bp) ellipse (27bp and 18bp);
  \draw (243bp,90bp) node {$\mathscr{N}_c$};
\end{scope}
  % Node: c1
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (99bp,90bp) ellipse (27bp and 18bp);
  \draw (99bp,90bp) node {$\mathscr{N}_b$};
\end{scope}
  % Node: c10
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (459bp,18bp) ellipse (27bp and 18bp);
  \draw (459bp,18bp) node {$\mathscr{N}_k$};
\end{scope}
  % Node: root
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (243bp,162bp) ellipse (27bp and 18bp);
  \draw (243bp,162bp) node {$\mathscr{N}_a$};
\end{scope}
  % Node: c6
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (171bp,18bp) ellipse (27bp and 18bp);
  \draw (171bp,18bp) node {$\mathscr{N}_g$};
\end{scope}
  % Node: c5
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (99bp,18bp) ellipse (27bp and 18bp);
  \draw (99bp,18bp) node {$\mathscr{N}_f$};
\end{scope}
  % Node: c4
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (27bp,18bp) ellipse (27bp and 18bp);
  \draw (27bp,18bp) node {$\mathscr{N}_e$};
\end{scope}
%
\draw[gray,thin,dashed] (0bp,58bp) -- (500bp,58bp);
\draw[gray,thin,dashed] (0bp,133bp) -- (500bp,133bp);

\draw (470bp,180bp) node[darkgray] {\scriptsize $s_a$};
\draw (470bp,122bp) node[darkgray] {\scriptsize $s_a - 1$};
\draw (470bp,47bp) node[darkgray] {\scriptsize $s_a - 2$};
\end{tikzpicture}