### 2014-11-13 bounds.tex

\documentclass[twoside,11pt]{article} % For LaTeX2e
\usepackage{jmlr2e}
\usepackage{hyperref}
\hypersetup{colorlinks,
            urlcolor=blue,
            citecolor=green,
            linkcolor=red}
\usepackage{url}
%\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09
\usepackage{mathrsfs}
%\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
%\usepackage{amssymb}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{color}

\usepackage{pstricks}
\usepackage{tikz}
\usepackage{subcaption}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

% Misc. math commands.  I couldn't figure out how to make the Unicode cat
% character, sadly.
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\var}{\operatorname{var}}

\newcommand{\K}{\mathcal{K}}
\newcommand{\hilbert}{\mathcal{H}}
\newcommand{\Real}{\mathcal{R}}

\renewcommand{\algorithmicforall}{\textbf{for each}}
\renewcommand{\algorithmicrepeat}{\textbf{do}}
\renewcommand{\algorithmicuntil}{\textbf{while}}

%\newtheorem{defn}{Definition}
%\newtheorem{thm}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newtheorem{cor}{Corollary}
\newtheorem{defn}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{cor}{Corollary}

\begin{document}

\title{Plug-and-play dual-tree algorithm runtime analysis}

\author{\name Ryan R. Curtin \email ryan@ratml.org\\
  \addr School of Computational Science and Engineering\\
  Georgia Institute of Technology\\
  Atlanta, GA 30332-0250, USA
  \AND
  \name Dongryeol Lee \email drselee@gmail.com\\
  \addr GE Research\\
  Schenedectadadaty, NY\\
  \AND
  \name William B. March \email march@ices.utexas.edu\\
  \addr Austin, TX, USA\\
  \AND
  \name Parikshit Ram \email p.ram@gatech.edu\\
  \addr Skytree
}

\editor{Unknown}

\maketitle

\begin{abstract}
%A large collection of machine learning algorithms contain subproblems that can
%be efficiently solved with trees.  In the situation where a dual-tree algorithm
%with cover trees is used to solve a problem, rigorous worst-case runtime bounds
%can often be shown.  A worst-case runtime bound is proven for all cover tree
%dual-tree algorithms; this bound can be employed to give useful runtime bounds
%for individual problems.  We show a new, tighter bound of $O(c^9 \nu N)$ for
%all-nearest-neighbor search and also provide linear runtime bounds for
%approximate kernel density estimation and exact range count/search under some
%mild assumptions.  These are the first linear runtime bounds we are aware of for
%range count/search.
Numerous machine learning algorithms contain pairwise statistical problems in
their core. Often, tree structures are used to solve these problems efficiently.
Dual-tree algorithms using cover trees have rigorous worst-case runtime
guarantees for some of these problems. In this paper, we present a {\em
problem-independent} runtime guarantee for {\em any} dual-tree algorithm using
the cover tree, separating out the problem-dependent and the problem-independent
elements. This allows us to just plug in bounds for the problem-dependent
elements to get runtime guarantees for dual-tree algorithms for any pairwise
statistical problem. We demonstrate this plug-and-play procedure for
nearest-neighbor search and approximate kernel density estimation to get
improved runtime guarantees.  Under some assumptions, we also present the first
linear runtime guarantee for dual-tree based range search.
\end{abstract}

\section{Dual-tree algorithms}

A surprising number of machine learning algorithms have computational
bottlenecks that can be expressed as pairwise statistical problems.
Nearest neighbor search is one such problem, naively requiring $O(N)$ time
to answer in single query in a reference set of size $N$; answering $O(N)$
queries subsequently requires prohibitive $O(N^2)$ time. Kernel density
estimation is also a pairwise statistical problem, naively requiring $O(N^2)$
time to answer $O(N)$ queries. The reference set is typically indexed with
spatial data structures to accelerate this type of computation
\citep{finkel1974quad, langford2006}; these result in $O(\log N)$ runtime per
query under favorable conditions.

\citet{nbody} generalized the fast multipole method from
computational physics to obtain dual-tree algorithms.  These are extremely
useful when there are large query sets, not just a few query points.  Instead of
building a tree on the reference set and searching with each query point
separately, Gray and Moore suggest also building a query tree and traversing
both the query and reference trees simultaneously (a {\it dual-tree traversal},
from which the class of algorithms takes its name).

Dual-tree algorithms can be easily understood through the recent framework of
\citet{curtin2013tree}: two trees (a query tree and a reference
tree) are traversed by a {\it pruning dual-tree traversal}.  This traversal
visits combinations of nodes from the trees in some sequence (each combination
consisting of a query node and a reference node), calling a problem-specific
\texttt{Score()} function to determine if the node combination can be pruned.
If not, then a problem-specific \texttt{BaseCase()} function is called for each
combination of points held in the query node and reference node.  This has
significant similarity to the more common single-tree branch-and-bound
algorithms, except that the algroithm must recurse into child nodes of {\it
both} the query tree and reference tree.

There exist numerous dual-tree algorithms for problems as diverse as kernel
density estimation \citep{gray2003nonparametric}, mean shift \citep{wang2007fast},
minimum spanning tree calculation \citep{march2010euclidean}, $n$-point
correlation function estimation \citep{march2012fast}, max-kernel search
\citep{curtin2013fast}, particle smoothing \citep{klaas2006fast}, variational
inference \citep{amizadeh2012variational}, and range search \citep{nbody}, to name
a few.

Some of these algorithms are derived using the cover tree \citep{langford2006}, a
data structure with beneficial theoretical qualities.  Dual-tree all
nearest-neighbor search and approximate kernel density estimation with cover
trees have $O(N)$ runtime guarantees for $O(N)$ queries \citep{ram2009}; minimum
spanning tree calculation scales as $O(N \log N)$ \citep{march2010euclidean}.
Other problems have similar worst-case guarantees \citep{curtin2014dual,
march2013multi}.

In this work we combine the generalization of \citet{curtin2013tree} with the
theoretical results of \citet{langford2006} and others in order to develop a
worst-case runtime bound for any dual-tree algorithm when the cover tree is
used.

Section \ref{sec:preliminaries} lays out the required background, notation, and
introduces the cover tree and its associated theoretical properties.  Readers
familiar with the cover tree literature and dual-tree algorithms
\citep[especially][]{curtin2013tree} may find that section to be review.
Following that, we introduce an intuitive measure of cover tree imbalance in
Section \ref{sec:imbalance}.  This measure of imbalance is then used to prove
the main result of the paper in Section \ref{sec:bound}, which is a worst-case
runtime bound for generalized dual-tree algorithms.  We apply this result to
three specific problems: nearest neighbor search (Section \ref{sec:nns}),
approximate kernel density estimation (Section \ref{sec:akde}), and range search
/ range count (Section \ref{sec:rs}), showing linear runtime bounds for each of
those algorithms.  Each of these bounds is an improvement on the
state-of-the-art, and in the case of range search, is the first such bound.  We
then investigate the empirical performance of cover trees for these algorithms,
with a focus on understanding the dataset-dependent constants that our bounds
use.

\section{Preliminaries}
\label{sec:preliminaries}

\input{bounds-files/sec01-supporting-lemmas.tex}

\section{Tree Imbalance}
\label{sec:imbalance}

\input{bounds-files/imbalance.tex}

\section{General runtime bound}
\label{sec:bound}

\input{bounds-files/general-runtime-bound.tex}

\section{Nearest neighbor search}
\label{sec:nns}

\input{bounds-files/sec02-nearest-neighbor-search.tex}

\section{Approximate kernel density estimation}
\label{sec:akde}

\input{bounds-files/sec03-approx-kde.tex}

\section{Range Search and Range Count}
\label{sec:rs}

\input{bounds-files/sec04-range-search.tex}

%\section{Range Search with Aspect Ratio}

%\input{bounds-files/sec05-range-search-aspect-ratio.tex}

\section{Conclusion}

\input{bounds-files/conclusion.tex}

%\bibliographystyle{plain}
\bibliography{bounds}

% Appendix doesn't have anything useful or non-straightforward.
\appendix
\input{appendix-ec}

\end{document}