### 2014-12-24 bounds-files/conclusion.tex
Ready for review.

We have presented a unified framework for bounding the runtimes of dual-tree
algorithms that use cover trees and the standard cover tree pruning dual-tree
traversal (Algorithm \ref{alg:cover-tree-dual}).  In order to produce an
understandable bound, we have introduced the notion of cover tree imbalance; one
possible interesting direction of future work is to empirically and
theoretically minimize this quantity by way of modified tree construction
algorithms; this is likely to provide both tighter runtime bounds and also
accelerated empirical results.

Our main result, Theorem \ref{thm:ct-runtime}, allows plug-and-play runtime
bounding of these algorithms.  We have shown that Theorem \ref{thm:ct-runtime}
is useful for bounding the runtime of nearest neighbor search (Theorem
\ref{thm:nns}), approximate kernel density estimation (Theorem
\ref{thm:kde-bound}), exact range count, and exact range search (Theorem
\ref{thm:rs}).  With our contribution, bounding a cover tree dual-tree algorithm
is straightforward and only involves bounding the maximum size of the reference
set, $| R^* |$.