### 2014-12-29 bounds-files/alternative_imbalance.tex
updated abstract and intro, typo fixes, added alternative imbalance notes

The purpose of this section is to explore an alternative to the current 
definition of the tree imbalance. I'll start with some intuition, then provide
an alternative definition. Finally, I'll try to re-write the proof of Theorem
1 using this definition.

I started by wondering if the tree imbalance Equation~\ref{eqn_node_imbalance}
can be bounded in terms of the expansion constant. I tried to construct 
examples where the expansion constant is small, but the tree imbalance is 
large.

The only one I could come up with is this: construct a set of $N/2$ points with
a small expansion constant (i.e. points on a lattice in low dimensions). Make a 
copy of this set, then place it arbitrarily far away from the first set.  The 
expansion constant of the combined set is no larger than for one of the small
sets. However, the gap in level between the root and its children is as 
large as we want.

The problem is that a dual-tree algorithm on this set will generally still be 
efficient. The key is that all of the levels between the root and second level 
of the tree don't matter. The algorithm can skip over all of them in one 
recursion.

So, the idea is to see if we can bound the runtime in terms of the number of 
missing levels \emph{which actually exist} somewhere in the tree. In the 
bichromatic case, we will need to take into account the levels of the other
tree.

Let $T_q$ and $T_r$ be a query and reference tree. Let $s_1$ and $s_2$ be two 
scales such that $s_1 > s_2$. 
Define $\mathcal{M}(s_1, s_2, T_q, T_r)$ to be the number of levels between 
$s_1$ and $s_2$ (inclusive) which one (or both) trees have a node at.

We provide the following definition of the imbalance of a node: 
\begin{definition}
  Let $T_r$ and $T_q$ be two cover trees. 
  Let $N$ be a node in either tree at scale $s$ with parent at scale $s_p$. 
  Then, define the imbalance of the node to be
  \begin{equation}
    \mathcal{I}(N) = \mathcal{M}(s_p, s, T_r, T_q)
  \end{equation}
  The imbalance of the root is defined to be $0$ and the level of a leaf is 
  defined to be $s_{\textrm{min}} - 1$ where $s_{\textrm{min}}$ is the minimum
  level of any non-leaf in either tree.

  The imbalance of a tree is the sum of the imbalances of all nodes.
\end{definition}

Note that the imbalance of a tree is $\bigO(N)$, since in the worst case 
each node has a different level from all of the others.

Is this really true, though? If we're counting for query nodes, then we might
count each reference node many times. This is where my thinking starts to break 
down. . .