### 2015-08-14 figures/cover-tree-ex.tex
Add an example cover tree figure.

%%
\begin{scope}
  \pgfsetstrokecolor{black}
  \definecolor{strokecol}{rgb}{1.0,1.0,1.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{1.0,1.0,1.0};
  \pgfsetfillcolor{fillcol}
  \filldraw (0bp,0bp) -- (0bp,252bp) -- (414bp,252bp) -- (414bp,0bp) -- cycle;
\end{scope}
\begin{scope}
  \pgfsetstrokecolor{black}
  \definecolor{strokecol}{rgb}{1.0,1.0,1.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{1.0,1.0,1.0};
  \pgfsetfillcolor{fillcol}
  \filldraw (0bp,0bp) -- (0bp,252bp) -- (414bp,252bp) -- (414bp,0bp) -- cycle;
\end{scope}
\begin{scope}
  \pgfsetstrokecolor{black}
  \definecolor{strokecol}{rgb}{1.0,1.0,1.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{1.0,1.0,1.0};
  \pgfsetfillcolor{fillcol}
  \filldraw (0bp,0bp) -- (0bp,252bp) -- (414bp,252bp) -- (414bp,0bp) -- cycle;
\end{scope}
  \pgfsetcolor{black}
  % Edge: ne -> ni
  \draw [->] (215.35bp,72.765bp) .. controls (219.71bp,64.283bp) and (225.15bp,53.714bp)  .. (234.7bp,35.147bp);
  % Edge: nd -> ng
  \draw [->] (99bp,71.697bp) .. controls (99bp,63.983bp) and (99bp,54.712bp)  .. (99bp,36.104bp);
  % Edge: nd -> nf
  \draw [->] (84.43bp,74.834bp) .. controls (74.25bp,64.938bp) and (60.476bp,51.546bp)  .. (41.796bp,33.385bp);
  % Edge: nb -> ne
  \draw [->] (207bp,143.7bp) .. controls (207bp,135.98bp) and (207bp,126.71bp)  .. (207bp,108.1bp);
  % Edge: nc -> nk
  \draw [->] (323.26bp,144.71bp) .. controls (335.67bp,120.23bp) and (359.24bp,73.739bp)  .. (378.62bp,35.537bp);
  % Edge: nb -> nd
  \draw [->] (188.19bp,148.81bp) .. controls (171bp,137.67bp) and (145.38bp,121.06bp)  .. (117.6bp,103.05bp);
  % Edge: na -> nb
  \draw [->] (264.43bp,218.83bp) .. controls (254.25bp,208.94bp) and (240.48bp,195.55bp)  .. (221.8bp,177.38bp);
  % Edge: na -> nc
  \draw [->] (287.35bp,216.76bp) .. controls (291.71bp,208.28bp) and (297.15bp,197.71bp)  .. (306.7bp,179.15bp);
  % Edge: nc -> nj
  \draw [->] (315bp,143.87bp) .. controls (315bp,119.67bp) and (315bp,75.211bp)  .. (315bp,36.189bp);
  % Edge: ne -> nh
  \draw [->] (198.65bp,72.765bp) .. controls (194.29bp,64.283bp) and (188.85bp,53.714bp)  .. (179.3bp,35.147bp);
  % Node: nh
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (171bp,18bp) ellipse (27bp and 18bp);
  \draw (171bp,18bp) node {$\mathscr{N}_2$};
\end{scope}
  % Node: ni
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (243bp,18bp) ellipse (27bp and 18bp);
  \draw (243bp,18bp) node {$\mathscr{N}_5$};
\end{scope}
  % Node: nj
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (315bp,18bp) ellipse (27bp and 18bp);
  \draw (315bp,18bp) node {$\mathscr{N}_1$};
\end{scope}
  % Node: nk
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (387bp,18bp) ellipse (27bp and 18bp);
  \draw (387bp,18bp) node {$\mathscr{N}_3$};
\end{scope}
  % Node: na
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (279bp,234bp) ellipse (27bp and 18bp);
  \draw (279bp,234bp) node {$\mathscr{N}_a$};
\end{scope}
  % Node: nb
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (207bp,162bp) ellipse (27bp and 18bp);
  \draw (207bp,162bp) node {$\mathscr{N}_b$};
\end{scope}
  % Node: nc
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (315bp,162bp) ellipse (27bp and 18bp);
  \draw (315bp,162bp) node {$\mathscr{N}_c$};
\end{scope}
  % Node: nd
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (99bp,90bp) ellipse (27bp and 18bp);
  \draw (99bp,90bp) node {$\mathscr{N}_d$};
\end{scope}
  % Node: ne
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (207bp,90bp) ellipse (27bp and 18bp);
  \draw (207bp,90bp) node {$\mathscr{N}_e$};
\end{scope}
  % Node: nf
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (27bp,18bp) ellipse (27bp and 18bp);
  \draw (27bp,18bp) node {$\mathscr{N}_0$};
\end{scope}
  % Node: ng
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \definecolor{fillcol}{rgb}{0.83,0.83,0.83};
  \pgfsetfillcolor{fillcol}
  \filldraw [opacity=1] (99bp,18bp) ellipse (27bp and 18bp);
  \draw (99bp,18bp) node {$\mathscr{N}_4$};
\end{scope}

\draw[gray,thin,dashed] (0bp,53bp) -- (420bp,53bp);
\draw[gray,thin,dashed] (0bp,125bp) -- (420bp,125bp);
\draw[gray,thin,dashed] (0bp,197bp) -- (420bp,197bp);

\draw (400bp,250bp) node[darkgray] {\scriptsize scale $1$};
\draw (400bp,185bp) node[darkgray] {\scriptsize scale $0$};
\draw (400bp,115bp) node[darkgray] {\scriptsize scale $-1$};
\draw (400bp,45bp) node[darkgray] {\scriptsize scale $-\infty$};

%