### 2015-08-14 figures/cover-tree-points.tex

\coordinate (p0) at (0, 0);
\coordinate (p1) at (1.0, -1.3);
\coordinate (p2) at (-0.3, 0.85);
\coordinate (p3) at (1.5, -1.8);
\coordinate (p4) at (0.4, -0.2);
\coordinate (p5) at (-0.6, 1.3);

\node [draw, circle, inner sep=1pt, fill] at (p0) { };
\node [below] at (p0) { $p_0$ };
\node [draw, circle, inner sep=1pt, fill] at (p1) { };
\node [below] at (p1) { $p_1$ };
\node [draw, circle, inner sep=1pt, fill] at (p2) { };
\node [below] at (p2) { $p_2$ };
\node [draw, circle, inner sep=1pt, fill] at (p3) { };
\node [below] at (p3) { $p_3$ };
\node [draw, circle, inner sep=1pt, fill] at (p4) { };
\node [below] at (p4) { $p_4$ };
\node [draw, circle, inner sep=1pt, fill] at (p5) { };
\node [below] at (p5) { $p_5$ };