### 2015-08-14 responses_to_reviewers.tex
Add some responses to the reviewers.

\documentclass[10pt]{article} % For LaTeX2e
\usepackage{url}
\usepackage{amsmath}
\usepackage{mathrsfs}

\begin{document}

Dear reviewers:

Thank you for your time in reviewing our manuscript.  We appreciate the detailed
comments and have incorporated several important changes:

\begin{itemize}
  \item The imbalance has been renamed $I_t(\cdot)$ (from $i_t(\cdot)$), to
avoid confusion with the index $i$ which used in other contexts.

  \item The spelling and grammatical errors have been fixed.

  \item The quantity $\theta$ is now defined outside Lemma 4.

  \item We have added a table of empirical results for the imbalance of cover
trees on a handful of real-world datasets (Table 1).

  \item We have added Figure 1, an example cover tree, which should help readers
develop intuition for cover trees.

  \item We have emphasized in the introduction that these bounds only apply to
cover trees, not other types of trees.  Unfortunately, other types of trees have
not (at this time) had any comparable adaptive analysis performed, so we cannot
compare our results with results for other tree types.
\end{itemize}

Now, we answer a few questions that came up during the reviews:

\begin{quote}
{\bf (Reviewer 1)} In particular, it seems non-trivial to derive similar bounds
for other efficient search structures like kd-trees, ball trees, or vantage
point trees, as these are not amenable to the notion of imbalancedness that the
authors use in their bounds. Are existing bounds for these tree structures
better or worse (or similar) than the bounds derived here? And more importantly,
what is the performance of cover trees compared to these other structures in
practice?
\end{quote}

We are not aware of existing worst-case bounds for other tree structures that
are non-trivial and very loose (i.e. $O(N^2)$ for nearest neighbor search), so
unfortunately we cannot compare our theoretical results in any meaningful way.
The empirical performance of cover trees is often comparable with that of
$kd$-trees, metric trees, and other types of trees.  Here is an interesting
empirical survey that supports this claim:

A.M. Kibriya, E. Frank, "An empirical comparison of exact nearest neighbour
algorithms", PKDD 2007.  (Table 1 is of particular interest, though it is
restricted to single-tree search.)
\url{http://www.cs.waikato.ac.nz/~ml/publications/2007/KibriyaAndFrankPKDD07.pdf}

In addition, our experiences with mlpack development seem to indicate that
$kd$-trees and cover trees can perform comparably.  Of course, the actual
performance is dependent on the data; sometimes cover trees are faster (usually
in medium dimensions), sometimes $kd$-trees are faster (usually in lower
dimensions).

We've added some references in the introduction to highlight the
already-documented empirical performance of dual-tree algorithms, too.

\begin{quote}
{\bf (Reviewer 1)} The authors state that "tree imbalance is near-linear with
the number of points", which would imply the run-time bounds behave more like
O(N*N) in practice?
\end{quote}

We think there has been a slight misunderstanding here.  The bound is of the
form O(N + imbalance), so linear imbalance scaling still gives O(N) scaling.

\begin{quote}
{\bf (Reviewer 2)} Does the result extend to k-nearest neighbor search?
\end{quote}

Yes; you get a bound of the form $O(k(N + I_t(\mathscr{T}_q) + \theta))$.
Essentially the only thing that needs to change is Equation 24, where the ball
$B_{S_r \cup \{ p_q \}}(p_q, d(p_q, p_r^*) / 2)$ now contains a maximum of $k$
points instead of $1$ point.

\begin{quote}
{\bf (Reviewer 3)} Lemma 1 proof: the definition of the expansion constant is an
inequality.  How can you assert the equality?
\end{quote}

This was a typo.  We have replaced the $=$ with $\le$ as it should be.

\begin{quote}
{\bf (Reviewer 3)} I find it strange that you are defining the "top" and
"minimum"... why not "top" and "bottom" or "maximum" and "minimum"?
\end{quote}

Unfortunately we are forced to use these terms here.  In our view, ``bottom
scale'' would imply the bottom of the tree, and the minimum scale is not the
scale of the bottom of the tree (that scale is $-\infty$).  We also can't use
the term ``maximum scale'', because that term is used heavily in the proof of
Lemma 2 to refer to the maximum scale present in the set $R$:

$$s_r^{\max} = \max_{\mathscr{N}_i \in R} s_i.$$

Again, thank you all for your comments!

\end{document}