### 2015-08-25 bounds-files/imbalance.tex
Update some comments about the experiment.

It is well-known that imbalance in trees leads to degradation in performance;
for instance, a $kd$-tree node with every descendant in its left child except
one is effectively useless.  A $kd$-tree full of nodes like this will perform
abysmally for nearest neighbor search, and it is not hard to generate a
pathological dataset that will cause a $kd$-tree of this sort.

This sort of imbalance applies to all types of trees, not just $kd$-trees.  In
our situation, we are interested in a better understanding of this imbalance for
cover trees, and thus endeavor to introduce a more formal measure of imbalance
which is correlated with tree performance.  Numerous measures of tree
imbalance have already been established; one example is that proposed by
\citet{colless1982review}, and another is Sackin's index \citep{sackin1972good},
but we aim to capture a different measure of imbalance that utilizes the leveled
structure of the cover tree.

We already know each node in a cover tree is indexed with an integer level (or
scale).  In the explicit representation of the cover tree, each non-leaf node
has children at a lower level.  But these children need not be strictly one
level lower; see Figure \ref{fig:imbalance}.  In Figure
\ref{fig:imbalance-good}, each cover tree node has children that are strictly
one level lower; we will refer to this as a {\em perfectly balanced cover tree}.
Figure \ref{fig:imbalance-bad}, on the other hand, contains the node
$\mathscr{N}_m$ which has two children with scale two less than $s_m$.  We will
refer to this as an {\em imbalanced cover tree}.  Note that in our definition,
the balance of a cover tree has nothing to do with differing number of
descendants in each child branch but instead only missing levels.

\begin{figure}
\begin{subfigure}[b]{0.585\textwidth}
  \begin{center}
    \input{figures/balanced-cover-tree.tex}
  \end{center}
  \caption{Balanced cover tree.}
  \label{fig:imbalance-good}
\end{subfigure}
\begin{subfigure}[b]{0.415\textwidth}
  \begin{center}
    \input{figures/imbalanced-cover-tree.tex}
  \end{center}
  \caption{Imbalanced cover tree.}
  \label{fig:imbalance-bad}
\end{subfigure}
\caption{Balanced and imbalanced cover trees.}
\label{fig:imbalance}
\end{figure}

An imbalanced cover tree can happen in practice, and in the worst cases, the
imbalance may be far worse than the simple graphs of Figure \ref{fig:imbalance}.
Consider a dataset with a single outlier which is very far away from all of the
other points\footnote{Note also that for an outlier sufficiently far away, the
expansion constant is $N - 1$, so we should expect poor performance with the
cover tree anyway.}.  Figure \ref{fig:outlier} shows what happens in
this situation: the root node has two children; one of these children has only
the outlier as a descendant, and the other child has the rest of the points in
the dataset as a descendant.  In fact, it is easy to find datasets with a
handful of outliers that give rise to a chain-like structure at the top of the
tree: see Figure \ref{fig:outliers} for an illustration\footnote{As a side note,
this behavior is not limited to cover trees, and can happen to mean-split
$kd$-trees too, especially in higher dimensions.}.

\begin{figure}
\begin{center}
  \input{figures/single-outlier-tree.tex}
\end{center}
\caption{Single-outlier cover tree.}
\label{fig:outlier}
\end{figure}

\begin{figure}
\begin{center}
  \input{figures/multiple-outlier-tree.tex}
\end{center}
\caption{A multiple-outlier cover tree.}
\label{fig:outliers}
\end{figure}

A tree that has this chain-like structure all the way down, which is similar to
the $kd$-tree example at the beginning of this section, is going to perform
horrendously; motivated by this observation, we define a measure of tree
imbalance.

\begin{defn}
The {\it cover node imbalance} $I_n(\mathscr{N}_i)$ for a cover tree node
$\mathscr{N}_i$ with scale $s_i$ in the cover tree $\mathscr{T}$ is defined as
the cumulative number of missing levels between the node and its parent
$\mathscr{N}_p$ (which has scale $s_p$).  If
the node is a leaf child (that is, $s_i = -\infty$), then number of missing
levels is defined as the difference between $s_p$ and $s_{\min} - 1$ where
$s_{\min}$ is the smallest scale of a non-leaf node in $\mathscr{T}$.  If
$\mathscr{N}_i$ is the root of the tree, then the cover node imbalance is 0.
Explicitly written, this calculation is

\begin{equation}
I_n(\mathscr{N}_i) = \begin{dcases*}
  s_p - s_i - 1 & if $\mathscr{N}_i$ is not a leaf and not the root node \\
  \max(s_p - s_{\min} - 1, \; 0) & if $\mathscr{N}_i$ is a leaf \\
  0 & if $\mathscr{N}_i$ is the root node.
  \end{dcases*}
  \label{eqn_node_imbalance}
\end{equation}
\end{defn}

This simple definition of cover node imbalance is easy to calculate, and using
it, we can generalize to a measure of imbalance for the full tree.

\begin{defn}
\label{def:imbalance}
The {\it cover tree imbalance} $I_t(\mathscr{T})$ for a cover tree $\mathscr{T}$
is defined as the cumulative number of missing levels in the tree.  This can be
expressed as a function of cover node imbalances easily:

\begin{equation}
I_t(\mathscr{T}) = \sum_{\mathscr{N}_i \in \mathscr{T}} I_n(\mathscr{N}_i).
\end{equation}
\end{defn}

A perfectly balanced cover tree $\mathscr{T}_b$ with no missing levels has
imbalance $I_t(\mathscr{T}_b) = 0$ (for instance, Figure
\ref{fig:imbalance-good}).  A worst-case cover tree $\mathscr{T}_w$ which is
entirely a chain-like structure with maximum scale $s_{\max}$ and minimum scale
$s_{\min}$ will have imbalance $I_t(\mathscr{T}_w) \sim N (s_{\max} -
s_{\min})$.  Because of this chain-like structure, each level has only one node
and thus there are at least $N$ levels; or, $s_{\max} - s_{\min} \ge N$, meaning
that in the worst case the imbalance is quadratic in $N$\footnote{Note that in
this situation, $c \sim N$ also.}.

However, for most real-world datasets with the cover tree implementation in {\bf
mlpack} \citep{mlpack2013} and the reference implementation
\citep{langford2006}, the tree imbalance is near-linear with the number of
points.  We have constructed cover trees on $N$ uniformly subsampled points
from a variety of datasets and calculated the imbalance; see Table
\ref{tab:imbalance} for the results.  These results are normalized with respect
to $N$, for which the values of $5000$, $50000$, and $500000$ were chosen.
The `power', `susy', `higgs', and `covertype' datasets are found in the UCI
Machine Learning Repository \cite{ucimlrepository}, the `mnist' dataset is from
\cite{lecun2000mnist}, and the `lcdm` and `sdss' datasets are Sloan Digital Sky
Survey data \cite{adelman2008sixth}.  The imbalances on each of these datasets
tend to be near-linear.

Currently, no cover tree construction algorithm specifically aims to minimize
imbalance.

\begin{table}[tb]
\begin{center}
\begin{tabular}{| c | c | c | c | c |}
\hline
 & & \multicolumn{3}{|c|}{Imbalance} \\
\hline
Dataset & $d$ & $N = 5k$ & $N = 50k$ & $N = 500k$ \\
\hline
%lcdm & 3 & 29402 & 313190 & 3400305 \\
lcdm & 3 & 5.88 & 6.26 & 6.80 \\
%sdss & 4 & 17691 & 191716 & 1483711 \\
sdss & 4 & 3.54 & 3.83 & 2.97 \\
% power & 7 & 22152 & 202050 & 1345825 \\
power & 7 & 4.43 & 4.04 & 2.69 \\
% susy & 18 & 3686 & 64231 & 930969 \\
susy & 18 & 0.74 & 1.28 & 1.86 \\
% higgs & 29 & 404 & 1272 & 73793 \\
higgs & 29 & 0.08 & 0.03 & 0.15 \\
% covertype & 55 & 9680 & 100344 & 1319447 \\
covertype & 55 & 1.94 & 2.01 & 2.64 \\
% mnist & 784 & 4004 & 79375 & 983554 \\
mnist & 784 & 0.80 & 1.59 & 1.97 \\
\hline
\end{tabular}
\end{center}
\caption{Empirically calculated tree imbalances, normalized by $N$.}
\label{tab:imbalance}
\end{table}

%To demonstrate this, Table \ref{tab:emp_imbalance} shows empirically
%calculated cover tree imbalance for cover trees on a number of datasets built
%with both the cover tree implementation in {\bf mlpack} \citep{mlpack2013} and
%the original reference implementation \citep{langford2006}.

%\begin{table}[tb]
%\begin{center}
%\begin{tabular}{|c|c|c|r|r|}
%\hline
%Dataset & $N$ & $d$ & reference & mlpack \\
%\hline
%wine & & & & \\
%isolet & & & & \\
%corel & & & & \\
%mnist & 70000 & 784 & & \\
%bio & & & & \\
%phy & & & & \\
%covertype & & & & \\
%LCDM & & & & \\
%\hline
%\end{tabular}
%\end{center}
%\caption{Empirical tree imbalance calculations.}
%\label{tab:emp_imbalance}
%\end{table}

%{\bf TODO: Can any bound be proven on $i_t(\mathscr{T})$?  I haven't thought of
%anything successfully yet.}