### 2015-10-31 submission-cover-letter.tex
Add submission cover letter.

\documentclass[twoside,11pt]{article}
\usepackage{jmlr2e}
\usepackage{hyperref}
\hypersetup{colorlinks, urlcolor=blue, citecolor=green, linkcolor=red}
\usepackage{url}

\usepackage{mathrsfs}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\begin{document}

Dear editors and reviewers: \\

\medskip

Thank you for your time and consideration in reviewing our paper,
``Plug-and-play dual-tree algorithm runtime analysis''.  We are glad to see that
it has been accepted, and we have addressed the issues pointed out by each
reviewer.  We apologize for the long delay in our final submission; multiple
authors have changed jobs since the original submission, and one (the
corresponding author) was unfortunately particularly busy defending his thesis.
With those things now firmly in the past, we found the time to complete our
revisions and assemble this document (which is a sort of changelog from the
submission).  We'll address the concerns of each reviewer individually in the
following sections.

\medskip

\section{Reviewer 1}

\begin{quote}{\it
In particular, it seems non-trivial to derive similar bounds for other
efficient search structures like kd-trees, ball trees, or vantage point trees,
as these are not amenable to the notion of imbalancedness that the authors use
in their bounds.
}\end{quote}

This is absolutely true, and we have re-emphasized at the end of Section 1 that
these results are only applicable to cover trees.  To our knowledge nobody has
been able to prove better than worst-case runtime bounds on $kd$-trees or ball
trees or any other type of trees for exact nearest neighbor search.  Part of the
reason can be seen in the complexity of the cover tree: very specific structure
and traversal types are required for bounding the number of reference nodes
considered at any point during the traversal.  With the notion of the expansion
constant, at least, it is unclear how to do this for any type of tree other than
the cover tree (or direct derivatives thereof).

However, these bounds do support the empirical observations that dual-tree
algorithms with other types of trees (like the aforementioned $kd$-tree or ball
tree) do tend to scale linearly.  We've mentioned this in footnote 1 at the
bottom of page 3, providing some useful references.

\begin{quote}{\it
And more importantly, what is the performance of cover trees compared to these
other structures in practice?
}\end{quote}

Performance of the cover tree vs. the $kd$-tree has been shown to be about
comparable in a survey by Kibriya et~al.: ``An Empirical Comparison of Exact
Nearest Neighbour Search Algorithms'', 2007.  However, that paper only considers
the single-tree nearest neighbor search algorithm, which isn't what we are
considering in this paper.  Our own experiences (which are readily verifiable
with mlpack's {\tt allknn} program) show that cover trees perform comparably to
$kd$-trees on most datasets, and for larger dimensionalities tend to outperform
$kd$-trees.

\begin{quote}{\it
Can the authors present some empirical results on the actual tree imbalance
measured on real data sets, to allow the reader to gauge the importance of the
imbalancedness value in the O(N * imbalance) runtimes?
}\end{quote}

To this end, we've added Table 2, on page 9, which has some empirical
simulations.

\begin{quote}{\it
The authors state that "tree imbalance is near-linear with the number of
points", which would imply the run-time bounds behave more like O(N*N) in
practice?
}\end{quote}

The bound is of the form $O(N + i(\mathscr{T}_q))$, not $O(N i(\mathscr{T}_q))$,
so linear (or better) imbalance implies $O(N)$ scaling, not $O(N^2)$.

\section{Reviewer 2}

\begin{quote}{\it
1) It would be nice to have a better understanding of the relationship between
the expansion constant and possible values of these tree-dependent quantities.
Is there a similar amortized run time bound of the form poly(c)n?  Are there
examples where these tree-related quantities can be constant in n, while is c is
not?  Having such a discussion would be helpful for the reader.
}\end{quote}

We agree, but unfortunately it is unclear how to relate the imbalance to the
expansion constant despite quite some time trying.  We have elaborated on the
intuition behind the imbalance somewhat, showing (and describing) some
worst-case situations.

\begin{quote}{\it
The authors consider 1-nearest neighbor search, finding the closest point to
each query point.  For this problem, the monochromatic case (where the reference
set is the query set) is not very meaningful, so the simplified bound after
Theorem 2 is not very sensible.
}\end{quote}

We've added a footnote, number 6 on page 19, which describes the {\tt
BaseCase()} modification necessary to perform meaningful monochromatic nearest
neighbor search.  We apologize for the confusion.

\begin{quote}{\it
Does the result extend to k-nearest neighbor search?
}\end{quote}

Yes, the result can be re-derived for $k$-nearest neighbor search, but it
becomes significantly more complicated.  The proof strategy is basically
identical; the major change will be the bound function $B(\mathscr{N}_q)$.  For
brevity we chose to omit this extension.

\begin{quote}{\it
The statement of Theorem 2 is missing a closing ")" after the imbalance
notation.
}\end{quote}

Thank you for pointing this out; this is now fixed.

\begin{quote}{\it
It's not a good idea to have definitions inside lemmas, so I'd strongly suggest
to define theta outside Lemma 4.  For one thing, Lemma 4 defines theta under the
condition that both $S_q$ and $S_r$ are have $O(N)$ points while Theorem 2 is
stated more generally.  Speaking of Lemma 4, $S_q$ and $S_r$ are defined as
sets, so you should probably have $|S_q| = |S_r| = O(N)$ in the statement of the
lemma.
}\end{quote}

This is a good suggestion and we have rewritten the affected lemma, adding a
definition for $\theta$.

\section{Reviewer 3}

\begin{quote}{\it
This naively requires $O(N)$ time to answer [in/a] single query
}\end{quote}

Thanks; fixed.

\begin{quote}{\it
I feel that an illustration of a cover tree would be very helpful.
}\end{quote}

We have added a detailed illustration in Figure 1.  We hope this is helpful.

\begin{quote}{\it
Lemma 1: "every point in $C$ is separated by" is not clear enough.  Do you mean
that no two points can be closer than $\delta$?  That every point in $C$ is
separated from every other point in $C$ by exactly $\delta$?  Spell out what you
mean.
}\end{quote}

We have clarified this; our meaning was ``every two points in $C$ are separated
by at least $\delta$''.

\begin{quote}{\it
Lemma 1 proof: ... The definition of the expansion constant is an inequality.
How can
you assert the equality?
}\end{quote}

This was a typo, and should have been an inequality.  This has been fixed;
thanks for pointing it out.

\begin{quote}{\it
You are already using "$i$" for the scale "$s_i$", so I would urge you to
use another symbol for imbalance.
}\end{quote}

We switched our notation to use capital $I$ as a result; thank you for the
suggestion.

\begin{quote}{\it
pg 10: "aspect ratio"... please indicate somehow that this is a
technical definition of "aspect ratio", not the colloquial one.
}\end{quote}

We've clarified that by aspect ratio we mean the maximum pairwise distance
divided by the minimum pairwise distance.

\begin{quote}{\it
pg 10:
in "The minimum scale fo the query tree", you have a subscript $T_r$, not $T_t$.
}\end{quote}

Thanks, this is now fixed.

\begin{quote}{\it
pg 10: I find it strange that you are defining the "top" and "minimum"... why
not "top" and "bottom" or "maximum" and "minimum"?
}\end{quote}

The goal here was to avoid reusing the term ``maximum scale'', because that
already refers to the maximum scale in a set, $s_r^{\max}$.  However, the term
``bottom'' would imply the scale of the leaves of the tree, but we actually want
the minimum non-leaf scale, so we believe ``minimum'' is a better term to use
here despite the strangeness of combining ``top'' and ``minimum''.

\begin{quote}{\it
eqn 10: just use normal parens on your min and max functions.  Typeset them
bigger to show grouping.
}\end{quote}

Fixed.

\begin{quote}{\it
I don't like the notion of a "poorly-behaved dataset".  The data are
the data!  It is the performance of the data structures \& algorithms
given that dataset that is poor.  Please rephrase.
}\end{quote}

This is an excellent point.  We have refactored our discussion for
clarification: ``These datasets for which trees perform poorly \ldots''.

\begin{quote}{\it
pg 14:
"Hence, the runtime of any dual-tree algorithm would be at least $O(N)$
using our bound, which matches the intuition that answering $O(N)$
queries would take at least $O(N)$ time."

You can't say anything about the minimum amount of time an algorithm
would take when all you have is a big-O bound.
}\end{quote}

We aren't saying that $O(N)$ is a lower bound on the amount of time necessary to
return results for $N$ queries---instead we are saying that when the various
parameters are plugged into our bound, it yields something linear (or greater)
in $N$, and this matches an intuitive inspection of these dual-tree algorithms.
We hope this clarifies our statement somewhat.

\begin{quote}{\it
pg 14:
Rework the paragraphs so that you don't need to start a paragraph with $\nu$.
}\end{quote}

Fixed.

\section{Final notes}

We have reworked a few other sections, provided some more explanation of the
dual-tree traversal and how it works, qualitatively, and fixed some other
minor grammatical errors that we missed for the original submission.  We have
now checked over our final draft thoroughly, addressed the reviewer comments,
and our final submission is enclosed.

Again, thanks to everyone involved in the review of this paper.  We hope our
revisions are satisfactory and again apologize for our high latency final
revision.  If there are any further issues, please let us know and we will not
hesitate to get them taken care of (this time with less turnaround time).

\end{document}